Modeling Intervention Strategies for United States TB Control

### 2013-08-20 in_progress/presentations/paper/draft.tex
made edits to paper draft

\subsection{Economic Modeling}
Further, estimates for the cost of active TB treatment and for the cost of
latent TB infection (LTBI) treatment were obtained from Dylan et. al. and So and
So adherence effectivenss. CITE ME!
Both these costs and any health state outcomes were discounted at a rate of 3\%
per year. From these estimated cost per treatment, total costs were obtained in
the following way. It was assumed that every patient in the US with active TB is
treated, not necessarily successfully, but further that every treatment costs
exactly the average cost per treatment. Thus, active TB costs could be tracked
simply by tracking the number of new active TB cases, and scaling by a
discounted cost per treatment rate. Given the structure of the model, further
granularity was obtained from the cost data by tracking separately the cost due
to new cases that stemmed from activation of a chronic LTBI case, or activation
of an acute LTBI case. These values give estimates for the US HCS TB cost due to
exogenous infection of TB vs the US HCS TB cost due to activation of LTBI. Each
of these components of the cost was a new differential equation tracked by the
model, also solved by $\texttt{lsoda}$. For LTBI treatment cost, treatment was
charged upon leaving the LTBI compartment due to the cumulative self-cure and
treatment rate given in the Hill Model. The fraction of these indviduals who
leave this compartment due to self cure was assumed to be zero. Given the
uncertainty in measurments of LTBI treatment cost, extensive sensitivity
analysis was performed on this parameter relative to final cost outcomes, which
showed that it had an XXXXXXXXX effect. In addition to US HCS cost, the system
was implemented so as to also track projected intervention implementation cost,
given user-inputted parameters relating to various possible intervention cost
strategies. The extended Hill Model was used to track the effect of many
interventions tested by the Hill Model as well as an additional intervention
strategy of curing cases of immigrating LTBI prior to entry. This intervention
strategy proved very promising and elimination year, final cost, and cost per
case averted were tracked for various levels of entering LTBI cure rate. The
interventions were implemented so as to take effect during the year 2013 and run
to 2100. The numerical DE solver $\texttt{lsoda}$ used was run with a time step
of $0.8$. Sensitivity analysis was performed on this parameter and reducing the
time step was shown to have minimal effect on final size or cost values.

@@

\subsection{Basic Population Breakdown}
The additional tracking capabilities offered several key insights into US TB
dynamics. Figure INSERT FIGURE HERE shows the yearly incidence of US TB, broken
down into infeciton source. It can be seen that the majority of the US TB load
is driven by activations of FB LTBI, followed by USB LTBI activations. This data
further agrees with the conclusions drawn by Hill, Becerra, and Castro about the
necessity of LTBI treatment in any valuable intervention strategy. Further,
figure INSERT FIGURE HERE shows a similarly sourced plot, but analyzing the
final US HCS costs due to TB. One can see that in this plot, roughly half of the
US TB HCS costs are due to activations of LTBI. Note that both of these plots
underestimate the impact LTBI activations play in the spread of TB, as every
LTBI activation to infectious TB contributes not only to incidence and costs
directly, but also indirectly by causing additional future cases, which is not
captured in these graphs. 
\subsubsection{Basic Reproduction Number}
The basic reproduction number of FB or USB cases of infectious TB was also
estimated by this system. From a theoretical perspective, we can think of the
total number of secondary infections over 100 years due to a FB or USB
infectious TB case as describing a geometric series in a large population.
Presuming there are no overlaps in infectious contacts, if a single case of
infectious TB in either population infects $p_f, p_u$ new cases in one year,
respectively, then we can say that over the course of 100 years, the total
number of cases infected will follow a geometric series. This analysis predicts
that over 100 years, one USB individual will infect $1.03$ other individuals,
whereas one FB individual will infect $.64$ other individuals. Experimentally,
these data were also analyzed, with results of $1.04$ and $.8$, respectively.

>>>>

\subsection{Economic Modeling}
Further, estimates for the cost of active TB treatment and for the cost of
latent TB infection (LTBI) treatment were obtained from Dylan et. al. and So and
So adherence effectivenss. CITE ME!
Both these costs and any health state outcomes were discounted at a rate of 3\%
per year. From these estimated cost per treatment, total costs were obtained in
the following way. It was assumed that every patient in the US with active TB is
treated, not necessarily successfully, but further that every treatment costs
exactly the average cost per treatment. Thus, active TB costs could be tracked
simply by tracking the number of new active TB cases, and scaling by a
discounted cost per treatment rate. Given the structure of the model, further
granularity was obtained from the cost data by tracking separately the cost due
to new cases that stemmed from activation of a chronic LTBI case, or activation
of an acute LTBI case. These values give estimates for the US HCS TB cost due to
exogenous infection of TB vs the US HCS TB cost due to activation of LTBI. Each
of these components of the cost was a new differential equation tracked by the
model, also solved by $\texttt{lsoda}$. For LTBI treatment cost, treatment was
charged upon leaving the LTBI compartment due to the cumulative self-cure and
treatment rate given in the Hill Model. The fraction of these indviduals who
leave this compartment due to self cure was assumed to be zero. Given the
uncertainty in measurments of LTBI treatment cost, extensive sensitivity
analysis was performed on this parameter relative to final cost outcomes, which
showed that it had an linear effect. In addition to US HCS cost, the system
was implemented so as to also track projected intervention implementation cost,
given user-inputted parameters relating to various possible intervention cost
strategies. The extended Hill Model was used to track the effect of many
interventions tested by the Hill Model as well as an additional intervention
strategy of curing cases of immigrating LTBI prior to entry. This intervention
strategy proved very promising and elimination year, final cost, and cost per
case averted were tracked for various levels of entering LTBI cure rate. The
interventions were implemented so as to take effect during the year 2013 and run
to 2100. The numerical DE solver $\texttt{lsoda}$ used was run with a time step
of $0.8$. Sensitivity analysis was performed on this parameter and reducing the
time step was shown to have minimal effect on final size or cost values.

@@

\subsection{Basic Population Breakdown}
The additional tracking capabilities offered several key insights into US TB
dynamics. Figure INSERT FIGURE HERE shows the yearly incidence of US TB, broken
down into infection source. It can be seen that the majority of the US TB load
is driven by activations of FB LTBI, followed by USB LTBI activations. This data
further agrees with the conclusions drawn by Hill, Becerra, and Castro about the
necessity of LTBI treatment in any valuable intervention strategy. Further,
figure INSERT FIGURE HERE shows a similarly sourced plot, but analyzing the
final US HCS costs due to TB. One can see that in this plot, roughly half of the
US TB HCS costs are due to activations of LTBI. Note that both of these plots
underestimate the impact LTBI activations play in the spread of TB, as every
LTBI activation to infectious TB contributes not only to incidence and costs
directly, but also indirectly by causing additional future cases, which is not
captured in these graphs. 
\subsubsection{Basic Reproduction Number}
The basic reproduction number of FB or USB cases of infectious TB was also
estimated by this system. From a theoretical perspective, we can think of the
total number of secondary infections over 100 years due to a FB or USB
infectious TB case as describing a geometric series in a large population.
Presuming there are no overlaps in infectious contacts, if a single case of
infectious TB in either population infects $p_f, p_u$ new cases in one year,
respectively, then we can say that over the course of 100 years, the total
number of cases infected will follow a geometric series. This analysis predicts
that over 100 years, one USB infection will lead to $1.03$ subsequent infections,
whereas one FB infection will lead to $.64$ subsequent infections. The full calculations
are included in the appendix.  Experimentally, these data were also analyzed, 
with results of $1.04$ and $.8$, respectively.
